---
title: "Dry Matter"
author: "Kolby Grint"
date: "2022-09-12"
output: pdf_document
---

```{r, echo = FALSE}
#Loading Packages
library(car)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(mgcv)
library(magrittr)

library(nlraa)
library(minpack.lm)
library(nlstools)
library(modelr)
library(devtools)
#source("eonr.R")
```


```{r}
#Accessing CSV Data
getwd()
data1= read.csv(file="2022Data.csv")

data1 = data1 %>%
  mutate(rep= as.factor(rep))

str(data1)
```

```{r}
data2 = data1 %>%
  group_by(N_trt, county) %>%
  summarise(DM_avg = mean(DM_total))

ggplot(data1, aes(x=N_trt, y= DM_total)) +
  #geom_point(data= data1, mapping = aes(y = DM_total, alpha = 0.01)) +
  geom_point(color= "red") + #red dots are just means
  facet_wrap(vars(county)) +
  theme(legend.position = "none")
```


# Dunn County
```{r}
data_Dunn <- data1 %>%
  filter(county == "Dunn") %>%
  slice(-c(10)) 
```

```{r}
data_Dunn2 = data_Dunn %>%
  group_by(N_trt) %>%
  summarise(DM_avg = mean(DM_total))

ggplot(data_Dunn2, aes(x=N_trt, y= DM_avg)) +
  geom_point(data = data_Dunn, mapping = aes(y = DM_total, alpha = 0.01)) +
  geom_point(color = "red") + #red dots are just means
  theme(legend.position = "none")
```


```{r}
lp <- function(N_trt, a, b, jp) {
  if_else(
    condition = N_trt < jp,
    true = a + b * N_trt,
    false = a + b * jp)
}
```

### Linear Plateau Method 1: Base R
```{r}
DunnLP1 <- nls(formula = DM_total ~ lp(N_trt, a, b, jp),
           data = data_Dunn,
           start = list(a = 1.45, #estimated y intercept
                        b = 0.002, #estimated slope
                        jp = 1.65)) #estimated join point on Y axis

#method failed, maybe the data is too messy
```
This function says that up until some join point, the relationship is linear, after which the stick breaks and it is a flat line (plateau). Sometimes this model is called broken-stick or linear-plus-plateau. That join point is important. In the context of soil fertility, it represents a critical soil test value or rate for the nutrient. The join point of a linear plateau will always be lower than with a quadratic plateau, an important consideration for another time and place.

### Linear Plateau Method 2
```{r}
DunnLP2 <- nlsLM(formula = DM_total ~ lp(N_trt, b0, b1, jp),
      data = data_Dunn,
      start = list(b0 = 1.45,
                   b1 = 0.002,
                   jp = 1.65))
#method worked. this one uses a different algorithm

summary(DunnLP2)
```

### Method 3 using a linear model to get better starting values
```{r}
#Using a linear model to get better starting values

Dunn.Linear <- lm(DM_total ~ N_trt, data = data_Dunn)
Dunn_start <-coef(Dunn.Linear)

#intercept was 1.43 and slope was 0.016
```

```{r}
DunnLP3 <- nlsLM(formula = DM_total ~ lp(N_trt, a, b, jp),
      data = data_Dunn,
      start = list(a = Dunn_start[1],
                   b = Dunn_start[2],
                   jp = median(data_Dunn$N_trt)))

summary(DunnLP3)
```

### Method 4 using a function that can self start
```{r}
DunnLP4 <- nlsLM(formula = DM_total ~ SSlinp(N_trt, a, b, jp), data = data_Dunn)

summary(DunnLP4)

```

```{r}
data_Dunn %>%
  ggplot(aes(N_trt, DM_total)) +
  geom_point() +
  geom_line(stat = "smooth", 
            method = "nlsLM", 
            formula = y ~ SSlinp(x, a, b, jp),
            se = FALSE,
            color= "red")
```

# Quadratic Plateau
The QP model is a type of segmented model, and QP is nice in that it has a curved component (important to biological systems) that meets a zero-slope plateau at the join point (important for researchers).

### Quadratic Plateau Function
```{r}
# a = intercept
# b = slope
# c = quadratic term (curvy bit)
# jp = join point = break point = critical concentration

qp <- function(N_trt, a, b, jp) {
    c <- -0.5 * b / jp
    if_else(condition = N_trt < jp,
            true  = a + (b * N_trt) + (c * N_trt * N_trt),
            false = a + (b * jp) + (c * jp * jp))
}
```

This function says that up until the join point, the relationship is a second-order polynomial (quadratic), after which it hits a flat line (plateau). This model is sometimes also called quadratic-plus-plateau. That join point is important. In the context of soil fertility, it represents a critical soil test concentration for the nutrient of interest. The join point of a “quad-plat” will always be higher than with a linear plateau, an important consideration for another time and place.


### Method 1: Base R
```{r}
DunnQP1 <- nls(formula = DM_total ~ qp(N_trt, a, b, jp),
            data = data_Dunn,
            start = list(a = 7.5, b = 0.03, jp = 12)) #didn't work

summary(fit1)
```

```{r}
quadratic.Dunn <- lm(DM_total ~ poly(N_trt, 2, raw = TRUE), data = data_Dunn)

start_values <- coef(quadratic.Dunn)

start_values
```

```{r}
DunnQP2 <- nls(formula = DM_total ~ qp(N_trt, a, b, jp),
            data = data_Dunn,
            start = list(a = start_values[1],
                         b = start_values[2],
                         jp = median(data_Dunn$N_trt)))

summary(DunnQP2)
```

### Method 2: nlraa
```{r}
DunnQP3 <- nls(formula = DM_total ~ SSquadp3xs(N_trt, a, b, xs), data= data_Dunn)

summary(DunnQP3)
```

```{r}
IC_tab(Dunn.Linear,DunnLP4,quadratic.Dunn,DunnQP3)
```
The linear model has the lowest AIC and BIC values.

# Plotting regression lines
```{r}
data_Dunn %>%
  ggplot(aes(N_trt, DM_total)) +
  geom_point() +
  geom_line(stat = "smooth", 
            method = "nlsLM", 
            formula = y ~ SSlinp(x, a, b, jp),
            se = FALSE,
            color= "blue") +
  geom_line(stat = "Smooth",
            method = "lm",
            color = "green") +
  geom_line(stat="smooth",
            method = "nls",
            formula = y ~ SSquadp3xs(x, a, b, jp),
            se = FALSE,
            color = "red") +
  labs(title = "Dunn County", x = "N-Rate", y = "wet biomass (ton/acre)")
```